Structuring College Access

### 2025-02-21 index.qmd
tweaks to presentation

- Student lists purchased by 14 public universities from 2016-2020
  - 414 student lists associated with 2.5 million prospects
  - Each prospect has a home zip code and a HS zip code (point geoms)
- Analyze purchases that include all Geomarkets in a metro area
  - Show who would be included/excluded had search criteria filtered on particular Geomarket(s)

@@

- Substantial inequality between and within Geomarkets
- In 1980, Black people are usually concentrated in poorest Geomarket
  - In some metros, this pattern persists (e.g., Cleveland)
  - In some metros, these Geomarkets experience gentrification

@@

- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic (and often Asian) students
- Poor, non-white Geomarkets produce disproportionate share of first-generation students
- Intersection of race and parental education
  - Asian students from low-income Geomarkets tend to be first-gen compared to Asain students from affluent Geomarkets
  - Same result for Hispanic students

@@

- Scholarship has not examined "b2b" quantification that classifies consumers for the sake of producers
- @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality

@@

- Scholarship focuses on schools and policies as "sorting machines" [@RN4873]
  - But EPS, Student Search Service products structure college access
  - College access structured by third-party vendors, products that amplify structural inequalities observed on demand-side
- Org field is "totality of relevant actors" [@RN527, p. 148]
  - Organizational field should be focus of regulation
  - Sociology should identify high-leverage actors/products in org field

>>>>

- Student lists purchased by 14 public universities from 2016-2020
  - 414 student lists associated with 2.5 million prospects
    - Data: search filter criteria; and de-identified student-level data
    - Each prospect has a home and a HS zip code (point geoms)
- Analyze purchases that include all Geomarkets in a metro area
  - Show who included/excluded if we filter on particular Geomarkets

@@

- Geomarkets highly correllated with race and class in many metros
- In 1980, Black people are usually concentrated in poorest Geomarket
  - In some metros, this pattern persists (e.g., Cleveland)

@@

- Poor, non-white Geomarkets produce disproportionate share of Black and Hispanic students
- Poor, non-white Geomarkets produce disproportionate share of first-generation students
- Intersection of race and parental education
  - Compared to students from affluent Geomarkets, Asian and Hispanic students from low-income Geomarkets tend to be first-gen
  - Targeting non-white students in affluent Geomarkets yields mostly non-first-gen

@@

- Hasn't examined quantification that classifies consumers for producers
  - Structural inequality $\rightarrow$ unequal student demand for college
  - College access structured by third-party products that encourage supply-side to amplify structural inequalities observed on demand-side
- @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality

@@

- Soc-of-ed focuses on schools (and policies) as "sorting machines"
  - But third-party vendors have structured college access for decades
- Org field is "totality of relevant actors" [@RN527, p. 148]
  - Organizational field should be focus of regulation
  - Sociology should identify high-leverage actors/products in org field
